---
title: "Asma Sultana"
output:
  html_document:
    df_print: paged
---



### Dataset load

 This project will reflect R programming and data science skills as a researcher.
```{r}


mortality = read.csv("Mortality.csv")
str(mortality)

```
### 4.Explain the dataset. 

##### In this load dataset, R programming is not showing  numerical column correctly as per as I guess. so generate those   column/feature my own thought.


 Numerical columns- age, creatinine_phosphokinase, ejection_fraction, platelets, serum_creatinine, serum_sodium, and time are numerical columns that likely represent   continuous variables.\
 Categorical columns: anaemia, diabetes, high_blood_pressure, sex, smoking are categorical columns that likely represent      binary or discrete variables.
 
 Target variable: The last column, DEATH_EVENT, is the target variable in the dataset, which likely indicates whether an individual experienced       a death event or not.

```{r}
numerical_columns = c("age", "creatinine_phosphokinase", "ejection_fraction", "platelets", "serum_creatinine", "serum_sodium", "time")
numerical_columns

categorical_columns = c("anaemia", "diabetes", "high_blood_pressure", "sex", "smoking")
categorical_columns

target_variable = "DEATH_EVENT"
target_variable

```

### 5.Correlation between Numerical feature
```{r}
## In this load dataset, R programming not showing numerical column correctly as per as I guess. so generate those  numerical feature my own thought. 

table = data.frame(

  a1= as.numeric(mortality$age),
  a2= as.numeric(mortality$creatinine_phosphokinase),
  a3= as.numeric(mortality$ejection_fraction),
  a4= as.numeric(mortality$platelets),
  a5= as.numeric(mortality$serum_creatinine),
  a6= as.numeric(mortality$serum_sodium),
  a7= as.numeric(mortality$time)
)  
table
```



#### Correlation numerical feature using loop
```{r}
 
## A named vector to store correlation coefficients
correlations = numeric(length = 6)

for (i in 2:7) {
  correlation = cor(table$a1, table[, i])
  correlations[i-1] = correlation #store
}

names(correlations) = c("a1-a2", "a1-a3", "a1-a4","a1-a5", "a1-a6", "a1-a7")

print(correlations)
```


```{r}
## Find out the strongest correlated feature and weakest correlated feature.

Max = (max(correlations))
cat(" Strongest correlation -> ", names(which.max(correlations)), ": ", Max, "\n") 

Min = (min(correlations))
cat(" Weakest correlation -> ", names(which.min(correlations)), ": ", Min, "\n") 



```
### 6.Subset for target feature
```{r}

subset0 = subset(mortality, DEATH_EVENT == 0)
subset1 = subset(mortality, DEATH_EVENT == 1)
subset0
subset1
```
#### Summary of all subsets 
```{r}
summary(subset0)
summary(subset1)

```

### Observations
  Comparing mean, median, min, and max values between subsets can reveal differences in variable distributions.
  Identifying patterns or trends specific to each subset may indicate factors influencing the target variable.
  Ensure balance of classes within subsets for better model performance.
  Further explore subsets using visualizations or additional analyses to understand their relationship with the target variable.




### 8. Determine correlation  for all the subsets

#### Correlation for subset0
```{r}

table2 = data.frame(

  a1= as.numeric(subset0$age),
  a2= as.numeric(subset0$creatinine_phosphokinase),
  a3= as.numeric(subset0$ejection_fraction),
  a4= as.numeric(subset0$platelets),
  a5= as.numeric(subset0$serum_creatinine),
  a6= as.numeric(subset0$serum_sodium),
  a7= as.numeric(subset0$time)
)  
table2


correlations = numeric(length = 6)

for (i in 2:7) {
  correlation = cor(table2$a1, table2[, i])
  correlations[i-1] = correlation #store
}

names(correlations) = c("a1-a2", "a1-a3", "a1-a4","a1-a5", "a1-a6", "a1-a7")

print(correlations)



Max0 = (max(correlations))
cat("\n Strongest correlation for subset0 -> ", names(which.max(correlations)), ": ", Max0, "\n") 

Min0 = (min(correlations))
cat(" Weakest correlation for subset0 -> ", names(which.min(correlations)), ": ", Min0, "\n") 


```

#### Correlation for subset0
```{r}

table3 = data.frame(

  a1= as.numeric(subset1$age),
  a2= as.numeric(subset1$creatinine_phosphokinase),
  a3= as.numeric(subset1$ejection_fraction),
  a4= as.numeric(subset1$platelets),
  a5= as.numeric(subset1$serum_creatinine),
  a6= as.numeric(subset1$serum_sodium),
  a7= as.numeric(subset1$time)
)  
table3


correlations = numeric(length = 6)

for (i in 2:7) {
  correlation = cor(table3$a1, table3[, i])
  correlations[i-1] = correlation #store
}

names(correlations) = c("a1-a2", "a1-a3", "a1-a4", "a1-a5", "a1-a6", "a1-a7")

print(correlations)



Max1 = (max(correlations))
cat("\n Strongest correlation for subset1 -> ", names(which.max(correlations)), ": ", Max1, "\n") 

Min1 = (min(correlations))
cat(" Weakest correlation for subset1 -> ", names(which.min(correlations)), ": ", Min1, "\n") 



```
### Observations
My observations are as follows as:
  a. In the overall dataset, the strongest correlation is between variables a1(age) and a5(serum_creatinine)) with a correlation coefficient of 0.1591871. The weakest correlation is between variables a1(age) and a7(time) with a correlation coefficient of -0.2240684.
  b. When considering subset0 of the data, the strongest correlation is still between variables a1(age) and a5(serum_creatinine)) with a correlation coefficient of 0.1375893. The weakest correlation within subset0 is between variables a1(age) and a4(platelets) with a correlation coefficient of -0.1086949. 
  c. For subset1 of the data, the strongest correlation is between variables a1(age) and a3(ejection_fraction) with a correlation coefficient of 0.216885. The weakest correlation within subset1 is between variables a1(age) and a7(time) with a correlation coefficient of -0.1876158.
